---
title: "Week 3 Homework"
author: "Nalin Kim"
date: '2022-09-25'
output: github_document
---

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(nycflights13)
theme_set(theme_bw())
```

#### 1. 
How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
sum(is.na(flights$dep_time))
map_dbl(flights, ~ sum(is.na(.x)))
```
There are 8255 flights with a missing `dep_time` variable, 8255 flights with a missing `dep_delay` variable, 8713 flights with a missing `arr_time` variable, 9430 with a missing `arr_delay` variable, and 9430 with a missing `air_time` variable. These missing variables represent that there are flights that failed to depart, failed to arrive, and canceled. On the other hand, it might be caused by possible data loss.

#### 2. 
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights %>%
  mutate(dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100),
       sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))
```
`%%` indicates x modulo y and `%/%` indicates integer division. The arithmetic operators mutate `dep_time` and `sched_dep_time` from actual times to number of minutes since midnight.

#### 3. 
Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()
```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(canceled = mean(is.na(dep_delay)),
            mean_dep_delay = mean(dep_delay, na.rm = T),
            mean_arr_delay = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = canceled)) +
  geom_point(aes(x =  mean_dep_delay, color = "Mean Departure Delay"), alpha = 0.5) +
  geom_smooth(aes(x =  mean_dep_delay, color = "Mean Departure Delay"), se = FALSE, method = "lm") + # regression line
  geom_point(aes(x = mean_arr_delay, color = "Mean Arrival Delay"), alpha = 0.5) +
  geom_smooth(aes(x = mean_arr_delay, color = "Mean Arrival Delay"), se = FALSE, method = "lm") + # regression line
  labs(x = "Mean Delay per Day", y = "Canceled Flights per Day", col = "Delay per Day")
```
The linear regression line shows a weak positive relationship between canceled flights and the average delay. In other words, there is a higher proportion of canceled flights per day when there is more average delay in minutes. It does not show a strong relationship, yet a few notable points show a robust association between cancellation and delay.